\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{../../styles/isaiah}
\usepackage{../../styles/components/isaiahOraclesOverview}

\begin{document}

\newpage
\begin{center}
{\Huge\bfseries Isaiah 19}
\end{center}
\vspace{10pt}

% Oracle Overview for Chapters 13-23
\isaiahOraclesOverview{5}

% Overview of Isaiah 19
\begin{overview}{Isaiah 19 — Overview}

\overviewsection[0]{%
\textbf{19:1-4: Oracle Introduction}
\begin{itemize}
    \item Judgment on Egypt's idols and civil war
\end{itemize}
}

\overviewsection[0]{%
\textbf{19:5-10: Economic Collapse}
\begin{itemize}
    \item Nile dries up, fishing and linen industries fail
\end{itemize}
}

\overviewsection[0]{%
\textbf{19:11-15: Wisdom Fails}
\begin{itemize}
    \item Egypt's counselors are foolish
\end{itemize}
}

\overviewsection[0]{%
\textbf{19:16-17: ``In that Day'' \#1}
\begin{itemize}
    \item Egypt will fear Judah
\end{itemize}
}

\overviewsection[1]{%
\textbf{19:18: ``In that Day'' \#2}
\begin{itemize}
    \item Five cities speak Hebrew and swear to the LORD
\end{itemize}
}

\overviewsection[1]{%
\textbf{19:19-22: ``In that Day'' \#3}
\begin{itemize}
    \item Altar to the LORD in Egypt, they will worship
\end{itemize}
}

\overviewsection[1]{%
\textbf{19:23-25: ``In that Day'' \#4}
\begin{itemize}
    \item Highway between Egypt and Assyria, blessing
\end{itemize}
}

\end{overview}

% Isaiah 19:1-11
\begin{biblicaloutline}[Isaiah 19:1-11]

\begin{versesection}{2em}
\versenum{1} An oracle concerning \highlightyellow{Egypt}.

Behold, the LORD is riding on a swift cloud
\poetryline and comes to \highlightyellow{Egypt};
and the \highlightpurple{idols} of \highlightyellow{Egypt} will tremble at his presence,
\poetryline and the heart of the \highlightyellow{Egyptians} will melt \highlightgreen{within them}.
\versenum{2} And I will stir up \highlightyellow{Egyptians} against \highlightyellow{Egyptians},
\poetryline and they will fight, each against another
\poetryline and each against his neighbor,
\poetryline city against city, kingdom against kingdom;
\versenum{3} and the \highlightblue{spirit} of the \highlightyellow{Egyptians} \highlightgreen{within them} will be emptied out,
\poetryline and I will confound their \highlightorange{counsel};
and they will inquire of the \highlightpurple{idols} and the sorcerers,
\poetryline and the mediums and the necromancers;
\versenum{4} and I will give over the \highlightyellow{Egyptians}
\poetryline into the hand of a hard master,
and a fierce king will rule over them,
\poetryline declares the \highlightred{Lord GOD of hosts}.

\versenum{5} And the waters of the sea will be dried up,
\poetryline and the river will be dry and parched,
\versenum{6} and its canals will become foul,
\poetryline and the branches of \highlightyellow{Egypt's} Nile will diminish and dry up,
\poetryline reeds and rushes will rot away.
\versenum{7} There will be bare places by the Nile,
\poetryline on the brink of the Nile,
and all that is sown by the Nile will be parched,
\poetryline will be driven away, and will be no more.
\versenum{8} The fishermen will mourn and lament,
\poetryline all who cast a hook in the Nile;
and they will languish
\poetryline who spread nets on the water.
\versenum{9} The \highlightpink{workers} in combed flax will be in despair,
\poetryline and the weavers of white cotton.
\versenum{10} Those who are the pillars of the land will be crushed,
\poetryline and all who \highlightgray{work} for pay will be grieved.

\versenum{11} The \highlightbrown{princes of Zoan} are utterly foolish;
\poetryline the wisest \highlightorange{counselors} of Pharaoh give stupid \highlightorange{counsel}.
How can you say to Pharaoh,
\poetryline "I am a son of the wise,
\poetryline a son of ancient kings"?
\end{versesection}

\end{biblicaloutline}

% Isaiah 19:12-25
\begin{biblicaloutline}[Isaiah 19:12-25]

\begin{versesection}{2em}
\versenum{12} Where then are your wise men?
\poetryline Let them tell you
\poetryline that they might \highlightlightblue{know} what the \highlightred{LORD of hosts} has \highlightorange{purposed} against \highlightyellow{Egypt}.
\versenum{13} The \highlightbrown{princes of Zoan} have become fools,
\poetryline and the princes of Memphis are deluded;
those who are the cornerstones of her tribes
\poetryline have made \highlightyellow{Egypt} stagger.
\versenum{14} The LORD has mingled \highlightgreen{within her} a \highlightblue{spirit} of confusion,
\poetryline and they will make \highlightyellow{Egypt} stagger in all its \highlightgray{deeds},
\poetryline as a drunken man staggers in his vomit.
\versenum{15} And there will be nothing for \highlightyellow{Egypt}
\poetryline that head or tail, palm branch or reed, may \highlightgray{do}.

\versenum{16} \highlightsilver{In that day} the \highlightyellow{Egyptians} will be like women, and tremble with \highlightdarkblue{fear} before the hand that the \highlightred{LORD of hosts} shakes over them. \versenum{17} And the land of Judah will become a terror to the \highlightyellow{Egyptians}. Everyone to whom it is mentioned will \highlightdarkblue{fear} because of the \highlightorange{purpose} that the \highlightred{LORD of hosts} has \highlightorange{purposed} against them.

\versenum{18} \highlightsilver{In that day} there will be five cities in the land of \highlightyellow{Egypt} that speak the language of Canaan and swear allegiance to the \highlightred{LORD of hosts}. One of these will be called the City of Destruction.

\versenum{19} \highlightsilver{In that day} there will be an altar to the LORD in the \highlightgreen{midst} of the land of \highlightyellow{Egypt}, and a pillar to the LORD at its border. \versenum{20} It will be a sign and a witness to the \highlightred{LORD of hosts} in the land of \highlightyellow{Egypt}. When they cry to the LORD because of oppressors, he will send them a savior and defender, and deliver them. \versenum{21} And the LORD will make himself \highlightlightblue{known} to the \highlightyellow{Egyptians}, and the \highlightyellow{Egyptians} will \highlightlightblue{know} the LORD \highlightsilver{in that day} and \highlightpink{worship} with sacrifice and offering, and they will make vows to the LORD and perform them. \versenum{22} And the LORD will strike \highlightyellow{Egypt}, striking and healing, and they will return to the LORD, and he will listen to their pleas for mercy and heal them.

\versenum{23} \highlightsilver{In that day} there will be a highway from \highlightyellow{Egypt} to Assyria, and Assyria will come into \highlightyellow{Egypt}, and \highlightyellow{Egypt} into Assyria, and the \highlightyellow{Egyptians} will \highlightpink{worship} with the Assyrians.

\versenum{24} \highlightsilver{In that day} Israel will be the third with \highlightyellow{Egypt} and Assyria, a blessing in the \highlightgreen{midst} of the earth, \versenum{25} whom the \highlightred{LORD of hosts} has blessed, saying, "Blessed be \highlightyellow{Egypt} my people, and Assyria the \highlightgray{work} of my hands, and Israel my inheritance."
\end{versesection}

\end{biblicaloutline}

\end{document}
